---
title: "Homework 3"
format: pdf
editor: visual
---

```{r, include=FALSE}
library(tidyverse)
library(palmerpenguins)

```

## Task 1
### a.
We are unable to use read.csv because it deals with data that is delimited by commas, not semicolons. Instead, read.csv2 is able to be used.

```{r}

read.csv2("C:/Users/ESLil/OneDrive/Desktop/ST 558/Repos/Homework3/Homework-3/Data/data.txt")

```
### b. 
```{r}

read.table("C:/Users/ESLil/OneDrive/Desktop/ST 558/Repos/Homework3/Homework-3/Data/data2.txt",
           skip = 1,
           sep = "6",
           col.names = c("x", "y", "z"),
           colClasses = c("factor", "numeric", "character")
           )
          

```

## Task 2
### a. 
```{r}

trailblazer <- read.csv("C:/Users/ESLil/OneDrive/Desktop/ST 558/Repos/Homework3/Homework-3/Data/trailblazer.csv")

glimpse(trailblazer)
```
### b. 
```{r}

trailblazer_longer <- trailblazer |>
  pivot_longer(
    cols = -Player,
    names_to = c("Game", "Location"),
    names_sep = "_",
    values_to = "Points"
  )

head(trailblazer_longer)

```
### c. 
```{r}
trailblazer_longer |>
  pivot_wider(names_from = Location, 
              values_from = Points) |>
  group_by(Player) |>
  summarize(mean_home = mean(Home, na.rm = T), 
            mean_away = mean (Away, na.rm = T)) |>
  mutate(Diff = mean_home - mean_away) |>
  arrange(Diff) 
```
The players that scored more points at home are the last 5 players in the tibble.

## Task 3
### a. 
<NULL> means that there are no bill length values for that observation. 

<dbl [52]> means that it is a numeric vector with 52 observations in the vector.

<list> means that each column is a list.

### b. 
```{r}

penguins |> 
  select(species, island) |>
  count(species, island) |>
  pivot_wider(names_from = island,
              values_fill = 0,
              values_from = n)
```
## Task 4

```{r}
penguins |>
  select(species, bill_length_mm) |>
  mutate(bill_length_mm = case_when(
    is.na(bill_length_mm) & species == "Adelie" ~ 26,
    is.na(bill_length_mm) & species == "Gentoo" ~ 30,
    TRUE ~ bill_length_mm
  )) |>
  arrange(bill_length_mm)
```

